Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-07-08 PAPER/simulation.tex
fixes for review

In this section we will test the proposed \disr{} approach to demonstrate
its effectiveness and compare it against topology agnostic approach
based on spanning trees and
broadcasting~\cite{Patwardhan05evaluatingthe}, to measure how \disr{}
performs in covering the network structure.

@@

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on a almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
tailored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of the \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yields
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links cannot be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node used from upper layer 
to inject the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{\emph{Node coverage}}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scales for increasing network sizes and defect rates
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

@@

In this section we analyze the results in terms of node
coverage and latency at different network sizes, defect rates and
bootstrap injection points.  In particular
Figure~\ref{fig:results_coverage} shows
node coverage for \disr{} and RPF tree based approach respectively. While
the first aim of \disr{} is not to reach the optimal coverage, we still
can observe a quite good performance as compared to the tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle.
For example Figure~\ref{fig:net} show a 30x30 network with a 25\%
defect in which the bottom-left part cannot be reached due
disconnected regions. Note that remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate do not introduce too much disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.70\textwidth]{pictures/net.ps}
\caption{Covered regions in a 30x30 network with 25\% defects}
\label{fig:net}
\end{figure*}

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demonstrating a working approach. 
These are optimizations are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming segment request,
the same request is cancelled after a given number of tries. This
gives to the preceding node receives a segment cancel and has the
chance of testing a different path instead of waiting indefinitely.
Default is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL}: when a segment request is cancelled
along a certain path (using a segment cancel), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times can desist, even if free links for going on are
available. This is not only a optimisation for shortening 
requests processes, but also solves some critical situations in which
request packets looped in a particular path without neither succeeding
nor failing.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}

>>>>

In this section we will test the proposed \disr{} approach to
demonstrate its effectiveness and compare it against a topology
agnostic approach based on spanning trees and
broadcasting~\cite{Patwardhan05evaluatingthe}, to measure how \disr{}
performs in covering the network structure. Note that a direct
comparison against the SR segment-based approach is not addressed
here, since how described in Section~\ref{sec:introduction} our
scenario assumes the non-feasibility of a centralized approach:
however, all the properties of the centralized approach should be
considered as preserved for all the nodes reached by the distributed
segment coverage of \disr{}. Thus, the main points that remain to be
addressed are:
\begin{itemize}
\item how \disr{} compares against state-of-art of tree based approach
applied to equivalent scenario
\item how \disr{} scales on large networks
\item the real feasibility of the required hardware on the limited
node size assumed
\end{itemize}

@@

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open-source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on an almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
tailored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yields
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links not able to be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node from upper layer that
injects the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{\emph{Node coverage}}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scale for increasing network sizes and defect rates.
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

@@

In this section we analyze the results in terms of node
coverage and latency with different network sizes, defect rates and
bootstrap injection points. In particular
Figure~\ref{fig:results_coverage} shows
node coverage for \disr{} and Reverse Path Forwarding (RPF) tree based approach respectively. While
the first aim of \disr{} is not to reach the optimal coverage, we still
can observe a quite good performance as compared to the tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle.
For example Figure~\ref{fig:net} shows a 30x30 network with a 25\%
defect in which the bottom-left part cannot be reached due to
disconnected regions. Note that the remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate does not introduce too many disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.70\textwidth]{pictures/net.ps}
\caption{Covered regions in a 30x30 network with 25\% of defects}
\label{fig:net}
\end{figure*}

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results, have been fixed to some good values and are not subject to
further investigation in this paper; once again the focus here is not the
optimal setup of segments, but just demonstrating a working approach. 
These parameters are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due to an incoming \texttt{SEGMENT\_REQUEST},
the request itself is cancelled after a given number of tries. This
gives to the preceding node on the path the
chance to test a different route instead of waiting indefinitely.
Default value is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL)}: when a segment request is cancelled
along a certain path (using a \texttt{SEGMENT\_CANCEL}), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times can desist, even if free links are
available. This is not only an optimization to shorten
requests' processes, but also solves some critical situations in which
request packets are blocked in routing loops in a particular path.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think about a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests have shown us that only
simulations using bootstrap nodes placed on edges would be heavily
affected by similar issues since these nodes start with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a single
defective link could prevent a starting segment packet from coming back to the
bootstrap node to close the loop and create the first segment.
\end{itemize}
